---
title: "Simulate Data"
author: "Lizzy Gibson"
date: "6/28/2019"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 999)
library(MNdata) 
library(reshape2) # melt()
library(compositions) # rlnorm.rplus()
library(MASS) # mvrnorm()
library(tidyverse)
library(gridExtra)
```

Simulate dataset with 1000 individuals and 35 chemicals to approx Mothers and Newborns cohort data.

## MN Data 

* 8 phenols, 9 phthalates, 10 pbdes, 8 pcbs
* phenols and phthalates are specific gravity adjusted
* all values < LOD are replaced by participant-specific LOD/sqrt(2)

```{r}
mn_data <- mn_edc %>% dplyr::select(1:52) %>%
  select_if(~n_distinct(.) > 1) %>% #drop column if all values are the same
  dplyr::select(1:18, pcb105, pcb74, pcb99, pcb118,
         pcb138_158, pcb153, pcb187, pcb180,
         grep("BDE", colnames(.))) %>%
  dplyr::select(-sid)

mn_data

cormat <- round(cor(mn_data, use = "complete.obs", method = c("spearman")),2)

melted_cormat <- melt(cormat) %>% rename(Correlation = value)

labelled <- melted_cormat %>% mutate(Class1 = ifelse(str_detect(Var1, "BDE"), "PBDEs",
                                                     ifelse(str_detect(Var1, "^M"), "Phthalates",
                                                            ifelse(str_detect(Var1, "pcb"), "PCBs",
                                                                   "Phenols")))) %>%
  mutate(Class2 = ifelse(str_detect(Var2, "BDE"), "PBDEs",
                         ifelse(str_detect(Var2, "^M"), "Phthalates",
                                ifelse(str_detect(Var2, "pcb"), "PCBs",
                                       "Phenols")))) %>%
  mutate(Class2 = factor(Class2, levels = c("Phthalates", "Phenols",
                                            "PCBs", "PBDEs")))

original <- ggplot(data = labelled, aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D",
                       midpoint = 0,
                       na.value = "transparent", limits = c(-1, 1)) +
  theme_grey(base_size = 15) + labs(x = "", y = "", title = "EDC correlations in Mothers & Newborns") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(),
        panel.spacing = unit(0, "lines"),
        strip.background = element_blank(),
        strip.placement = "outside", legend.position = "bottom",
        strip.text.x = element_text(size = 18),
        strip.text.y = element_text(size = 18)) +
  facet_grid(Class2 ~ Class1, scales = "free", switch = "both")
original

mn_data %>% 
  gather(key = variable) %>% 
  ggplot(aes(x = value)) + geom_density() + facet_wrap(.~variable, scales = "free")
```

## Simulations

### Multivariate log normal 

```{r}
# Vector of means in MN cohort
mean_log <- as_vector(map_df(log(mn_data), ~mean(., na.rm = TRUE)))

# Covariance matrix from MN cohort
var_log <- cov(log(mn_data), use = "complete.obs")
# missing values are handled by casewise deletion (and if there are no complete cases, that gives an error)

# Simulate with multivariate log normal function
set.seed(1988)
sim_mlvn <- rlnorm.rplus(1000, meanlog = mean_log, varlog = var_log) %>% as.matrix() %>% as_tibble()
colnames(sim_mlvn) <- names(mn_data)

sim_mlvn
```

### Multivariate normal function exponentiated

```{r}
set.seed(1988)
sim_exp <- exp(mvrnorm(n = 1000, mu = mean_log, Sigma = var_log)) %>% as.matrix() %>% as_tibble()
colnames(sim_exp) <- names(mn_data)

sim_exp

#write_csv(simulated, "./sim_data.csv")
```

## Visualize

```{r}
sim_cormat <- round(cor(sim_mlvn, use = "pairwise.complete.obs", method = c("spearman")),2)

sim_melted_cormat <- melt(sim_cormat) %>% rename(Correlation = value)

sim_labelled <- sim_melted_cormat %>% mutate(Class1 = ifelse(str_detect(Var1, "BDE"), "PBDEs",
                                                     ifelse(str_detect(Var1, "^M"), "Phthalates",
                                                            ifelse(str_detect(Var1, "pcb"), "PCBs",
                                                                   "Phenols")))) %>%
  mutate(Class2 = ifelse(str_detect(Var2, "BDE"), "PBDEs",
                         ifelse(str_detect(Var2, "^M"), "Phthalates",
                                ifelse(str_detect(Var2, "pcb"), "PCBs",
                                       "Phenols")))) %>%
  mutate(Class2 = factor(Class2, levels = c("Phthalates", "Phenols",
                                            "PCBs", "PBDEs")))

sim1_plot <- ggplot(data = sim_labelled, aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D",
                       midpoint = 0,
                       na.value = "transparent", limits = c(-1, 1)) +
  theme_grey(base_size = 15) + labs(x = "", y = "", title = "Simulated multivariate log normal") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(),
        panel.spacing = unit(0, "lines"),
        strip.background = element_blank(),
        strip.placement = "outside", legend.position = "bottom",
        strip.text.x = element_text(size = 18),
        strip.text.y = element_text(size = 18)) +
  facet_grid(Class2 ~ Class1, scales = "free", switch = "both")

#####################

sim2_cormat <- round(cor(sim_exp, use = "pairwise.complete.obs", method = c("spearman")),2)

sim2_melted_cormat <- melt(sim2_cormat) %>% rename(Correlation = value)

sim2_labelled <- sim2_melted_cormat %>% mutate(Class1 = ifelse(str_detect(Var1, "BDE"), "PBDEs",
                                                             ifelse(str_detect(Var1, "^M"), "Phthalates",
                                                                    ifelse(str_detect(Var1, "pcb"), "PCBs",
                                                                           "Phenols")))) %>%
  mutate(Class2 = ifelse(str_detect(Var2, "BDE"), "PBDEs",
                         ifelse(str_detect(Var2, "^M"), "Phthalates",
                                ifelse(str_detect(Var2, "pcb"), "PCBs",
                                       "Phenols")))) %>%
  mutate(Class2 = factor(Class2, levels = c("Phthalates", "Phenols",
                                            "PCBs", "PBDEs")))

sim2_plot <- ggplot(data = sim_labelled, aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D",
                       midpoint = 0,
                       na.value = "transparent", limits = c(-1, 1)) +
  theme_grey(base_size = 15) + labs(x = "", y = "", title = "Simulated exp(multivariate normal)") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(),
        panel.spacing = unit(0, "lines"),
        strip.background = element_blank(),
        strip.placement = "outside", legend.position = "bottom",
        strip.text.x = element_text(size = 18),
        strip.text.y = element_text(size = 18)) +
  facet_grid(Class2 ~ Class1, scales = "free", switch = "both")

original
sim1_plot
sim2_plot
```
