---
title: "Loop Simulations"
author: "Lizzy Gibson"
date: "3/5/2020"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
options(scipen = 999)
library(reshape2) # melt
library(MASS) # mvrnorm
library(tidyverse)
library(GGally)
```

## 1: Distinct Patterns

### Simulate Patterns

* Generate 4 non-negative vectors of length 50
* Pattern matrix is 4*50, patterns by chemicals
* Draw each pattern: number 1-2 if pattern is ON, basically 0 otherwise

```{r}
## Construct a matrix for 4 TRUE exposure patterns
## replace runif between 0--0.2 with rep 0
create_patterns_dist <- function (seed) {
  set.seed(seed)
  patterns_dist <- 
    rbind(c(runif(12, 1, 2),   rep(0, 38)),
          c(rep(0, 12),        runif(12, 1, 2),   rep(0, 26)),
          c(rep(0, 24),                           runif(13, 1, 2),  rep(0, 13)),
          c(rep(0, 37), runif(13, 1, 2)))
  patterns_dist
}

patterns_dist_iter <- tibble(seed = 1:100) %>% 
 mutate(true_patterns = map(seed, create_patterns_dist))
```

```{r}
patterns_dist_iter$true_patterns[[1]] %>% as_tibble() %>% 
  mutate(Pattern = 1:4) %>% 
  dplyr::select(Pattern, everything()) %>% 
  gather(key = Chemicals, value = Loadings, -Pattern) %>% 
  mutate(Chemicals = fct_inorder(Chemicals)) %>%
  ggplot(aes(x = Chemicals, y = Loadings)) + geom_col() +
  facet_wrap(.~Pattern) + theme_bw() +
  theme(axis.text.x = element_blank(),
        strip.background = element_blank()) +
  labs(title = "Distinct Patterns")
```

### Simulate Scores

* Generate non-negative scores
* 1,000 individuals with scores on each of 4 patterns
* Individual distributions are independent
* Patterns (columns) are NOT standardized
* Each pattern = log-normal distribution 
    * Mean = 1
    * Var approx 1
    * Cov = 0
   
```{r}
create_scores_dist <- function (seed) {
  scores_dist <- matrix(nrow = 1000, ncol = 4)
  set.seed(seed)
  scores_dist <- exp(mvrnorm(n = 1000, mu = rep(1, 4), # mu = rnorm(4, mean = 1), 
                          Sigma = diag(rep(1, 4)))) # Independent log normal dist
  scores_dist
}

scores_dist_iter <- patterns_dist_iter %>% 
 mutate(true_scores = map(seed, create_scores_dist))
```

### Simulate Chemical Exposures

* Matrix multiple scores and pattern loadings

```{r}
sim_dist <- scores_dist_iter %>% 
  mutate(chem = map2(true_scores, true_patterns, `%*%`))
```

### Simulate Noise

* Add noise
* Noise distribution Normal(0, s)
    * where s = simulated variance/10 for each variable
    
```{r}
add_noise <- function (seed, chemx) {
  noise <- list(mean = 0, var = apply(chemx, 2, var)/4)
  #noise <- list(mean = 0, var = rep(100, 50))
  set.seed(seed)
  chem_dist <- pmax(chemx + 
                      mvrnorm(n = 1000, mu = rep(0, 50), 
                                        Sigma = diag(noise$var)), 0)	
  chem_dist
}

sim_dist <- sim_dist %>% 
 mutate(sim = map2(seed, chem, add_noise))
```  

```{r}
dist <- sim_dist$sim[[2]]
melt(cor(dist, method = "spearman")) %>% 
  as_tibble() %>% 
  rename(Correlation = value) %>% 
  mutate(Var1 = as.factor(Var1),
         Var2 = as.factor(Var2)) %>% 
  ggplot(aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D",
                       midpoint = 0,
                       na.value = "transparent", limits = c(-1, 1)) +
  theme_classic() + labs(x = "", y = "") +
  theme(axis.text = element_blank(),
        legend.position = "bottom") + 
  ggtitle("Correlation w/ Distinct Underlying Patterns")
```

## 2: Overlapping Patterns

### Simulate Patterns

* Generate 4 non-negative vectors of length 50
* Pattern matrix is 4*50, patterns by chemicals
* Draw each pattern: number 1-2 if pattern is ON, 0.5-1 is the overlapping group, basically 0 otherwise
```{r}
# replace runif 0-0.2 with rep 0
create_patterns_over <- function (seed) {
  set.seed(seed)
  patterns_cor <-
    rbind(c(runif(12, 1, 2),   runif(12, 0.5, 1), rep(0, 26)),
          c(rep(0, 12),        runif(12, 1, 2),   runif(13, 0.5, 1), rep(0, 13)),
          c(rep(0, 24),                           runif(13, 1, 2),   runif(13, 0.5, 1)),
          c(runif(12, 0.5, 1), rep(0, 25),        runif(13, 1, 2)))
  patterns_cor
}

patterns_over_iter <- tibble(seed = 1:100) %>% 
 mutate(true_patterns = map(seed, create_patterns_over))
```

```{r}
patterns_over_iter$true_patterns[[1]] %>% as_tibble() %>% 
  mutate(Pattern = 1:4) %>% 
  dplyr::select(Pattern, everything()) %>% 
  gather(key = Chemicals, value = Loadings, -Pattern) %>% 
  mutate(Chemicals = fct_inorder(Chemicals)) %>%
  ggplot(aes(x = Chemicals, y = Loadings)) + geom_col() +
  facet_wrap(.~Pattern) + theme_bw() +
  theme(axis.text.x = element_blank(),
        strip.background = element_blank()) +
  labs(title = "Overlapping Patterns")
```

### Simulate Scores

Use distinct scores.

* Generate non-negative scores
* 1,000 individuals with scores on each of 4 patterns
* Individual distributions are independent
* Patterns (columns) are NOT standardized
* Each pattern = log-normal distribution 
    * Mean = 1
    * Var = 1
    * Cov = 0

### Simulate Chemical Exposures

* Matrix multiple scores and pattern loadings

```{r}
sim_over <- scores_dist_iter %>% 
  dplyr::select(seed, true_scores) %>% 
  full_join(., patterns_over_iter) %>% 
  mutate(chem = map2(true_scores, true_patterns, `%*%`))
```

### Simulate Noise

* Add noise
* Noise distribution Normal(0, s)
    * where s = simulated variance/4 for each variable
    
```{r}
sim_over <- sim_over %>% 
 mutate(sim = map2(seed, chem, add_noise))
```  

```{r}
over <- sim_over$sim[[1]]

melt(cor(over, method = "spearman")) %>% 
  as_tibble() %>% 
  rename(Correlation = value) %>% 
  mutate(Var1 = as.factor(Var1),
         Var2 = as.factor(Var2)) %>% 
  ggplot(aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", high = "#F8766D") +
  theme_classic() + labs(x = "", y = "") +
  theme(axis.text = element_blank(),
        legend.position = "bottom") + 
  ggtitle("Correlation w/ Overlapping Underlying Patterns")
```

## 3. Correlated Scores

### Simulate Scores

```{r}
create_scores_cor <- function (seed) {
  scores_cor <- matrix(nrow = 1000, ncol = 4)
  cov <- matrix( c(1,    0.25, 0.15, 0.15,
                  0.25, 1,    0.15, 0.15,
                  0.15, 0.15, 1,    0.25,
                  0.15, 0.15, 0.25, 1), nrow = 4)
  
  set.seed(seed)
  scores_cor <- exp(mvrnorm(n = 1000, mu = rep(1, 4), 
                          Sigma = cov)) # dists are kind of DEPENDENT
  }

scores_cor_iter <- patterns_over_iter %>% 
 mutate(true_scores = map(seed, create_scores_cor))
```

### Simulate Chemical Exposures

Multiple scores and vectors, and add to get the mixture. Use simulated overlapping patterns.

```{r}
sim_cor <- scores_cor_iter %>% 
  dplyr::select(seed, true_scores) %>% 
  full_join(., patterns_over_iter) %>% 
  mutate(chem = map2(true_scores, true_patterns, `%*%`))
```

### Simulate Noise

```{r}
sim_cor <- sim_cor %>% 
 mutate(sim = map2(seed, chem, add_noise))
```  

```{r}
simm <- sim_cor$sim[[1]]

ggcorr(as_tibble(simm), limits = FALSE,
       hjust = 0.85, size = 2, layout.exp = 1) +
  labs(x = "Simulated Correlation Matrix") +
  theme_minimal(base_size = 10)
```

## Save

```{r} 
sim_noise <- rbind(sim_dist %>% mutate(data = "Distinct"),
      sim_over %>% mutate(data = "Overlapping"),
      sim_cor %>% mutate(data = "Correlated"))
  
sim_noise

# for (i in 1:nrow(sim_dgp_rep1)) {
#   write_csv(as_tibble(sim_dgp_rep1$sim[[i]]), paste0("./Sims/dgp_rep1/sim_dgp_rep1_", i, ".csv"))
# }

save(sim_noise, file = "./Results/Simulations/sim_noise.RDA")
```

