---
title: "Loop Simulations"
author: "Lizzy Gibson"
date: "3/5/2020"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
options(scipen = 999)
library(reshape2) # melt
library(gridExtra) # gridarrange
library(bindata) # multivariate binomial dist
library(NMF) # rmatrix
library(MASS) # mvrnorm
library(tidyverse)
library(R.matlab)
library(MCMCpack) # for dirichlet
```

## 1: Distinct Patterns

### Simulate Patterns

Generate 4 non-negative vectors of length 50 “by hand” — vectors should be overlapping a bit, and at this stage don’t have to look too much like pollution. just make 20 (of 50) elements 1 and the rest 0.

Pattern matrix is 4*50, patterns by chemicals.

```{r}
create_patterns_dist <- function (seed) {
  set.seed(seed)
  patterns_dist <- rbind(rdirichlet(1, c(rep(10, 12), rep(0.1, 38))),
                         rdirichlet(1, c(rep(0.1, 12), rep(10, 12), rep(0.1, 26))),
                         rdirichlet(1, c(rep(0.1, 26), rep(10, 12), rep(0.1, 12))),
                         rdirichlet(1, c(rep(0.1, 36), rep(10, 14))))
  patterns_dist
}

patterns_dist_iter <- tibble(seed = 1:100) %>% 
 mutate(true_patterns = map(seed, create_patterns_dist))
```

```{r}
patterns_dist_iter[1,2][[1]][[1]] %>% as_tibble() %>% 
  mutate(Pattern = 1:4) %>% 
  dplyr::select(Pattern, everything()) %>% 
  gather(key = Chemicals, value = Loadings, -Pattern) %>% 
  mutate(Chemicals = fct_inorder(Chemicals)) %>%
  ggplot(aes(x = Chemicals, y = Loadings)) + geom_col() +
  facet_wrap(.~Pattern) + theme_bw() +
  theme(axis.text.x = element_text(angle = 45),
        strip.background = element_blank()) +
  labs(title = "Distinct Patterns")
```

### Simulate Scores

```{r}
create_scores_dist <- function (seed) {
  scores_dist <- matrix(nrow = 1000, ncol = 4)
  set.seed(seed)
  scores_dist <- exp(mvrnorm(n = 1000, mu = rep(0, 4), 
                          Sigma = diag(rep(1, 4)))) # independent log normal dist
  scores_dist <- apply(scores_dist, 2, function (x) x/sd(x)) # divide columns by standard dev 
  scores_dist
  }

scores_dist_iter <- patterns_dist_iter %>% 
 mutate(true_scores = map(seed, create_scores_dist))
```

### Simulate Chemical Exposures

Multiple scores and vectors, and add to get the mixture.

```{r}
sim_dist <- scores_dist_iter %>% 
  mutate(chem = map2(true_scores, true_patterns, `%*%`))
```

### Simulate Noise

```{r}
add_noise <- function (seed, chem) {
  noise <- list(mean=0, sd=1)
  set.seed(seed)
  chem_dist <- pmax(chem + rmatrix(1000, 50, dist=rnorm, mean=noise$mean, sd=noise$sd), 0)	
  chem_dist
}

sim_dist <- sim_dist %>% 
 mutate(sim = map2(seed, chem, add_noise))

apply(sim_dist[1,5][[1]][[1]]*10, 2, var)
```

```{r}
melt(cormat) %>% rename(Correlation = value) %>% 
ggplot(aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D",
                       midpoint = 0,
                       na.value = "transparent", limits = c(-1, 1)) +
  theme_grey() + labs(x = "", y = "") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.text = element_blank(),
        panel.spacing = unit(0, "lines"),
        strip.background = element_blank(),
        strip.placement = "outside", legend.position = "bottom") + ggtitle("Correlation w/ Distinct Underlying Patterns")
```

## 2: Correlated Patterns

### Simulate Patterns

```{r}
## Construct a matrix for 4 TRUE exposure patterns
create_patterns_cor <- function (seed) {
  set.seed(seed)
  patterns_cor <- rbind( rdirichlet(1, c(rep(10, 12),  rep(5, 12),   rep(0.1, 26))),
                         rdirichlet(1, c(rep(0.1, 12), rep(10, 12),  rep(5, 12),  rep(0.1, 14))),
                         rdirichlet(1, c(rep(0.1, 12), rep(5, 12),   rep(10, 12), rep(0.1, 14))),
                         rdirichlet(1, c(rep(5, 12),   rep(0.1, 24), rep(10, 14))))
  patterns_cor
}

patterns_cor_iter <- tibble(seed = 1:100) %>% 
 mutate(true_patterns = map(seed, create_patterns_cor))
```

```{r}
patterns_cor_iter[1,2][[1]][[1]] %>% as_tibble() %>% 
  mutate(Pattern = 1:4) %>% dplyr::select(Pattern, everything()) %>% 
  gather(key = Chemicals, value = Loadings, -Pattern) %>% 
  mutate(Chemicals = fct_inorder(Chemicals)) %>%
  ggplot(aes(x = Chemicals, y = Loadings)) + geom_col() +
  facet_wrap(.~Pattern) + theme_bw() +
  theme(axis.text.x = element_text(angle = 45),
        strip.background = element_blank()) +
  labs(title = "Overlapping Patterns")
```

### Simulate Scores
```{r}
create_scores_cor <- function (seed) {
  scores_cor <- matrix(nrow = 1000, ncol = 4)
  cov <- matrix( c(1,    0.15, 0.15, 0.15,
                  0.15, 1,    0.15, 0.15,
                  0.15, 0.15, 1,    0.15,
                  0.15, 0.15, 0.15, 1), nrow = 4)
  
  set.seed(seed)
  scores_cor <- exp(mvrnorm(n = 1000, mu = rep(1, 4), 
                          Sigma = cov)) # dists are kind of DEPENDENT
  }

scores_cor_iter <- patterns_cor_iter %>% 
 mutate(true_scores = map(seed, create_scores_cor))
```

### Simulate Chemical Exposures

Multiple scores and vectors, and add to get the mixture.

```{r}
sim_cor <- scores_cor_iter %>% 
  mutate(chem = map2(true_scores, true_patterns, `%*%`))
```

### Simulate Noise

```{r}
sim_cor <- sim_cor %>% 
 mutate(sim = map2(seed, chem, add_noise))
```  

```{r}
cormat2 <- round(cor(sim_cor[1,5][[1]][[1]]),2)

melt(cormat2) %>% rename(Correlation = value) %>% 
  ggplot(aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D",
                       midpoint = 0,
                       na.value = "transparent", limits = c(-1, 1)) +
  theme_grey() + labs(x = "", y = "") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.text = element_blank(),
        panel.spacing = unit(0, "lines"),
        strip.background = element_blank(),
        strip.placement = "outside", legend.position = "bottom") + ggtitle("Overlapping Patterns & Correlated Scores")
```

## 3: Overlapping Patterns

### Simulate Patterns

Use correlated patterns.

### Simulate Scores

Use distinct scores.

### Simulate Chemical Exposures

Multiple scores and vectors, and add to get the mixture.

```{r}
sim_over <- scores_dist_iter %>% 
  select(seed, true_scores) %>% 
  full_join(., scores_cor_iter %>% select(seed, true_patterns), by = "seed") %>% 
  mutate(chem = map2(true_scores, true_patterns, `%*%`))
```

### Simulate Noise

```{r}
sim_over <- sim_over %>% 
 mutate(sim = map2(seed, chem, add_noise))
```  

```{r}
cormat3 <- round(cor(sim_over[100,4][[1]][[1]]),2)
melt(cormat3) %>% rename(Correlation = value) %>% 
  ggplot(aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D",
                       midpoint = 0,
                       na.value = "transparent") +
  theme_grey() + labs(x = "", y = "") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.text = element_blank(),
        panel.spacing = unit(0, "lines"),
        strip.background = element_blank(),
        legend.position = "bottom") + 
        labs(title = "Overlapping patterns & independent scores")
```

### Save

```{r} 
sim_norm <- rbind(sim_dist %>% mutate(data = "Distinct"),
      sim_over %>% mutate(data = "Overlapping")) %>% 
  mutate(sim_type = "L1 Norm") %>%
 mutate(sim_var10 = map(sim, function(x) x*5)) %>% 
  pivot_longer(c(sim, sim_var10),
               names_to = "var",
               values_to = "sim") %>% 
  mutate(var = ifelse(var == "sim", "reg", "var10"))

sim_norm

#save(sim_norm, file = "./R/Sims/Iterate/sim_norm.RDA")

# for (i in 1:100) {
#   write_csv(as_tibble(sim_over[i,4][[1]][[1]]), paste0("./R/Sims/Iterate/sim_over_", i, "_stand.csv"))
# }
```
