---
title: "NMF & Sim Data -- Iterate"
author: "Lizzy Gibson"
date: "2/25/2020"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
options(scipen = 999)
library(MNdata) # my data
library(reshape2) # melt
library(gridExtra) # gridarrange
library(bindata) # multivariate binomial dist
library(NMF) # rmatrix
library(MASS) # mvrnorm
library(tidyverse)
library(R.matlab)
```

## 1: Distinct Patterns

### Simulate Patterns

Generate 4 non-negative vectors of length 50 “by hand” — vectors should be overlapping a bit, and at this stage don’t have to look too much like pollution. just make 20 (of 50) elements 1 and the rest 0.

Pattern matrix is 4*50, patterns by chemicals.

```{r}
## Construct a matrix for 4 TRUE exposure patterns
labels <-  c("pcb28", "pcb66", "pcb74", "pcb99", "pcb105", "pcb118", "pcb138_158", "pcb146", "pcb153", "pcb156", "pcb167",
          "pcb170", "pcb178", "pcb183", "pcb187", "pcb180", "pcb189", "pcb194", "pcb196_203", "pcb199", "pcb206", "pcb209",
          "BDE17", "BDE28", "BDE47", "BDE66", "BDE85", "BDE99", "BDE100", "BDE153", "BDE154", "BDE183", "BDE209", "MECPP",
          "MEHHP", "MEOHP", "MCPP", "MIBP", "MBP", "MBZP", "MEP", "MEHP", "dcp_24", "dcp_25", "b_pb", "bp_3", "m_pb",
          "p_pb", "tcs", "bpa")

create_patterns <- function (seed) {
  set.seed(seed)
  patterns_dist <- rbind(c(runif(22, 1, 2), rep(0, 11), rep(0, 17)),
                       c(rep(0, 22), runif(11, 1, 2), rep(0, 17)),
                       c(rep(0, 33), runif(9, 1, 2), rep(0, 8)),
                       c(rep(0, 33), rep(0, 9), runif(8, 1, 2)))
  colnames(patterns_dist) <- labels
  patterns_dist
}

patterns_dist_iter <- tibble(seed = 1:100) %>% 
 mutate(patterns = map(seed, ~create_patterns(seed = .x)))

patterns_dist_iter
```

### Simulate Scores

```{r}

create_scores <- function (seed) {
  scores_dist <- matrix(nrow = 1000, ncol = 4)
  set.seed(seed)
  scores_dist <- exp(mvrnorm(n = 1000, mu = rep(1, 4), 
                          Sigma = diag(rep(1, 4)))) # Independent log normal dist
  scores_dist
}

scores_dist_iter <- tibble(seed = 1:100) %>% 
 mutate(scores = map(seed, ~create_scores(seed = .x)))

scores_dist_iter
```

### Simulate Chemical Exposures

Multiple scores and vectors, and add to get the mixture.

```{r}
sim_dist0 <- full_join(scores_dist_iter, patterns_dist_iter, by = "seed")

sim_dist0 <- sim_dist0 %>% 
  mutate(chem = map2(scores, patterns, `%*%`))
```

### Simulate Noise

```{r}

add_noise <- function (chem) {
  noise <- list(mean=0, sd=1)
  set.seed(1988)
  chem_dist <- pmax(chem + rmatrix(1000, 50, dist=rnorm, mean=noise$mean, sd=noise$sd), 0)	
  chem_dist
}

sim_dist <- sim_dist0 %>% 
 mutate(sim = map(chem, ~add_noise(chem = .x)))

sim_dist
```  

### Visualize

```{r, echo = FALSE, fig.height = 7}
chem_dist <- sim_dist$sim[[30]] %>% 
  as_tibble(.)

cormat <- round(cor(chem_dist, use = "complete.obs"),2)

melted_cormat <- melt(cormat) %>% rename(Correlation = value)

ggplot(data = melted_cormat, aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D",
                       midpoint = 0,
                       na.value = "transparent", limits = c(-1, 1)) +
  theme_grey() + labs(x = "", y = "") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.text.x = element_blank(),
        panel.spacing = unit(0, "lines"),
        strip.background = element_blank(),
        strip.placement = "outside", legend.position = "bottom")
```

## 2: Overlapping Patterns

### Simulate Patterns

```{r}
## Construct a matrix for 4 TRUE exposure patterns
labels <-  c("pcb28", "pcb66", "pcb74", "pcb99", "pcb105", "pcb118", "pcb138_158", "pcb146", "pcb153", "pcb156", "pcb167",
          "pcb170", "pcb178", "pcb183", "pcb187", "pcb180", "pcb189", "pcb194", "pcb196_203", "pcb199", "pcb206", "pcb209",
          "BDE17", "BDE28", "BDE47", "BDE66", "BDE85", "BDE99", "BDE100", "BDE153", "BDE154", "BDE183", "BDE209", "MECPP",
          "MEHHP", "MEOHP", "MCPP", "MIBP", "MBP", "MBZP", "MEP", "MEHP", "dcp_24", "dcp_25", "b_pb", "bp_3", "m_pb",
          "p_pb", "tcs", "bpa")

create_patterns_over <- function (seed) {
  set.seed(seed)
  patterns_over <- rbind(c(runif(22, 1, 2),   runif(11, 0.5, 1), rep(0, 17)),
                         c(rep(0, 22),        runif(11, 1, 2),   runif(9, 0.5, 1),  rep(0, 8)),
                         c(rep(0, 33),                           runif(9, 1, 2),    runif(8, 0.5, 1)),
                         c(runif(22, 0.5, 1), rep(0, 11),        rep(0, 9),         runif(8, 1, 2)))
  
  colnames(patterns_over) <- labels
  patterns_over
}

patterns_over_iter <- tibble(seed = 1:100) %>% 
 mutate(patterns = map(seed, ~create_patterns_over(seed = .x)))

patterns_over_iter
```

### Simulate Scores

```{r}

create_scores_over <- function (seed) {
  
  scores_over <- matrix(nrow = 1000, ncol = 4)
  cov <- matrix( c(1,    0.25, 0.15, 0.15,
                  0.25, 1,    0.15, 0.15,
                  0.15, 0.15, 1,    0.25,
                  0.15, 0.15, 0.25, 1), nrow = 4)
  
  set.seed(seed)
  scores_over <- exp(mvrnorm(n = 1000, mu = rep(1, 4), 
                          Sigma = cov)) # dists are kind of DEPENDENT
  }

scores_over_iter <- tibble(seed = 1:100) %>% 
 mutate(scores = map(seed, ~create_scores_over(seed = .x)))

scores_over_iter
```

### Simulate Chemical Exposures

Multiple scores and vectors, and add to get the mixture.

```{r}
sim_over0 <- full_join(scores_over_iter, patterns_over_iter, by = "seed")

sim_over0 <- sim_over0 %>% 
  mutate(chem = map2(scores, patterns, `%*%`))
```

### Simulate Noise

```{r}
sim_over <- sim_over0 %>% 
 mutate(sim = map(chem, ~add_noise(chem = .x)))

sim_over
```  

### Visualize

```{r, echo = FALSE, fig.height = 7}
chem_over <- sim_over$sim[[5]] %>% 
  as_tibble(.)

cormat <- round(cor(chem_over, use = "complete.obs"),2)

melted_cormat <- melt(cormat) %>% rename(Correlation = value)

ggplot(data = melted_cormat, aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D",
                       midpoint = 0,
                       na.value = "transparent", limits = c(-1, 1)) +
  theme_grey() + labs(x = "", y = "") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.text.x = element_blank(),
        panel.spacing = unit(0, "lines"),
        strip.background = element_blank(),
        strip.placement = "outside", legend.position = "bottom")
```


