---
title: "Loop Simulations"
author: "Lizzy Gibson"
date: "3/5/2020"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
options(scipen = 999)
library(MNdata) # my data
library(reshape2) # melt
library(gridExtra) # gridarrange
library(bindata) # multivariate binomial dist
library(NMF) # rmatrix
library(MASS) # mvrnorm
library(tidyverse)
library(R.matlab)
```

## 1: Distinct Patterns

### Simulate Patterns

Generate 4 non-negative vectors of length 50 “by hand” — vectors should be overlapping a bit, and at this stage don’t have to look too much like pollution. just make 20 (of 50) elements 1 and the rest 0.

Pattern matrix is 4*50, patterns by chemicals.

```{r}
## Construct a matrix for 4 TRUE exposure patterns
labels <-  c("pcb28", "pcb66", "pcb74", "pcb99", "pcb105", "pcb118", "pcb138_158", "pcb146", "pcb153", "pcb156", "pcb167",
          "pcb170", "pcb178", "pcb183", "pcb187", "pcb180", "pcb189", "pcb194", "pcb196_203", "pcb199", "pcb206", "pcb209",
          "BDE17", "BDE28", "BDE47", "BDE66", "BDE85", "BDE99", "BDE100", "BDE153", "BDE154", "BDE183", "BDE209", "MECPP",
          "MEHHP", "MEOHP", "MCPP", "MIBP", "MBP", "MBZP", "MEP", "MEHP", "dcp_24", "dcp_25", "b_pb", "bp_3", "m_pb",
          "p_pb", "tcs", "bpa")

create_patterns_dist <- function (seed) {
  set.seed(seed)
  patterns_dist <- rbind(c(runif(22, 1, 2), rep(0, 11), rep(0, 17)),
                       c(rep(0, 22), runif(11, 1, 2), rep(0, 17)),
                       c(rep(0, 33), runif(9, 1, 2), rep(0, 8)),
                       c(rep(0, 33), rep(0, 9), runif(8, 1, 2)))
  colnames(patterns_dist) <- labels
  patterns_dist
}

patterns_dist_iter <- tibble(seed = 1:100) %>% 
 mutate(patterns = map(seed, ~create_patterns_dist(seed = .x)))
```

```{r, fig.height=5, fig.width=5}
patterns_dist_iter[1,2][[1]][[1]] %>% as_tibble() %>% 
  mutate(Pattern = 1:4) %>% 
  dplyr::select(Pattern, everything()) %>% 
  gather(key = Chemicals, value = Loadings, -Pattern) %>% 
  mutate(Chemicals = fct_inorder(Chemicals)) %>% 
    mutate(Group = ifelse(str_detect(Chemicals, "BDE"), "PBDEs",
                                                     ifelse(str_detect(Chemicals, "^M"), "Phthalates",
                                                            ifelse(str_detect(Chemicals, "pcb"), "PCBs",
                                                                   "Phenols")))) %>% 
  ggplot(aes(x = Chemicals, y = Loadings)) + geom_col(aes(fill = Group)) +
  facet_wrap(.~Pattern) + theme_bw() +
  theme(axis.text.x = element_blank(),
        strip.background = element_blank()) +
  labs(title = "Distinct Patterns")
```

### Simulate Scores

```{r}
create_scores_dist <- function (seed) {
  scores_dist <- matrix(nrow = 1000, ncol = 4)
  set.seed(seed)
  scores_dist <- exp(mvrnorm(n = 1000, mu = rep(1, 4), 
                          Sigma = diag(rep(1, 4)))) # Independent log normal dist
  scores_dist
}

scores_dist_iter <- tibble(seed = 1:100) %>% 
 mutate(scores = map(seed, ~create_scores_dist(seed = .x)))
```

### Simulate Chemical Exposures

Multiple scores and vectors, and add to get the mixture.

```{r}
sim_dist0 <- full_join(scores_dist_iter, patterns_dist_iter, by = "seed")

sim_dist0 <- sim_dist0 %>% 
  mutate(chem = map2(scores, patterns, `%*%`))
```

### Simulate Noise

```{r}
add_noise <- function (seed, chem) {
  noise <- list(mean=0, sd=1)
  set.seed(seed)
  chem_dist <- pmax(chem + rmatrix(1000, 50, dist=rnorm, mean=noise$mean, sd=noise$sd), 0)	
  chem_dist
  sim_noise <- chem_dist - chem
  list(chem_dist = chem_dist, noise = sim_noise)
}

sim_dist <- sim_dist0 %>% 
 mutate(sim = map2(seed, chem, ~add_noise(seed = .x, chem = .y)[[1]]),
        noise = map2(seed, chem, function(x, y) {add_noise(x, y)[[2]]}),
        sim = map(sim, function(x) scale(x, center = FALSE, scale = apply(x, 2, sd))))
```  

```{r}
dim(sim_dist[1,5][[1]][[1]])
cormat <- round(cor(sim_dist[2,5][[1]][[1]]),2)

melted_cormat <- melt(cormat) %>% rename(Correlation = value)

labelled <- melted_cormat %>% mutate(Class1 = ifelse(str_detect(Var1, "BDE"), "PBDEs",
                                                     ifelse(str_detect(Var1, "^M"), "Phthalates",
                                                            ifelse(str_detect(Var1, "pcb"), "PCBs",
                                                                   "Phenols")))) %>%
  mutate(Class2 = ifelse(str_detect(Var2, "BDE"), "PBDEs",
                         ifelse(str_detect(Var2, "^M"), "Phthalates",
                                ifelse(str_detect(Var2, "pcb"), "PCBs",
                                       "Phenols")))) %>%
  mutate(Class2 = factor(Class2, levels = c("Phthalates", "Phenols",
                                            "PCBs", "PBDEs")))

ggplot(data = labelled, aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D",
                       midpoint = 0,
                       na.value = "transparent", limits = c(-1, 1)) +
  theme_grey() + labs(x = "", y = "") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.text = element_blank(),
        panel.spacing = unit(0, "lines"),
        strip.background = element_blank(),
        strip.placement = "outside", legend.position = "bottom") + ggtitle("Correlation w/ Distinct Underlying Patterns")
```

### Save

```{r}
sim_dist <- sim_dist %>% rename(true_patterns = patterns, true_scores = scores) %>% dplyr::select(-chem, -noise)
         
#save(sim_dist, file = "./R/Sims/Iterate/sim_dist.RDA")

# for (i in 1:100) {
#   write_csv(as_tibble(sim_dist[i,4][[1]][[1]]), paste0("./R/Sims/Iterate/sim_dist_", i, "_stand.csv"))
# }
```

## 2: Correlated Patterns

### Simulate Patterns

```{r}
## Construct a matrix for 4 TRUE exposure patterns
create_patterns_cor <- function (seed) {
  set.seed(seed)
  patterns_cor <- rbind(c(runif(22, 1, 2),   runif(11, 0.5, 1), rep(0, 17)),
                         c(rep(0, 22),        runif(11, 1, 2),   runif(9, 0.5, 1),  rep(0, 8)),
                         c(rep(0, 33),                           runif(9, 1, 2),    runif(8, 0.5, 1)),
                         c(runif(22, 0.5, 1), rep(0, 11),        rep(0, 9),         runif(8, 1, 2)))
  
  colnames(patterns_cor) <- labels
  patterns_cor
}

patterns_cor_iter <- tibble(seed = 1:100) %>% 
 mutate(patterns = map(seed, ~create_patterns_cor(seed = .x)))
```

```{r, fig.width=5, fig.height=5}
patterns_cor_iter[1,2][[1]][[1]] %>% as_tibble() %>% 
  mutate(Pattern = 1:4) %>% dplyr::select(Pattern, everything()) %>% 
  gather(key = Chemicals, value = Loadings, -Pattern) %>% 
  mutate(Chemicals = fct_inorder(Chemicals)) %>% 
    mutate(Group = ifelse(str_detect(Chemicals, "BDE"), "PBDEs",
                                                     ifelse(str_detect(Chemicals, "^M"), "Phthalates",
                                                            ifelse(str_detect(Chemicals, "pcb"), "PCBs",
                                                                   "Phenols")))) %>% 
  ggplot(aes(x = Chemicals, y = Loadings)) + geom_col(aes(fill = Group)) +
  facet_wrap(.~Pattern) + theme_bw() +
  theme(axis.text.x = element_blank(),
        strip.background = element_blank()) +
  labs(title = "Overlapping Patterns")
```

### Simulate Scores

```{r}
create_scores_cor <- function (seed) {
  scores_cor <- matrix(nrow = 1000, ncol = 4)
  cov <- matrix( c(1,    0.25, 0.15, 0.15,
                  0.25, 1,    0.15, 0.15,
                  0.15, 0.15, 1,    0.25,
                  0.15, 0.15, 0.25, 1), nrow = 4)
  
  set.seed(seed)
  scores_cor <- exp(mvrnorm(n = 1000, mu = rep(1, 4), 
                          Sigma = cov)) # dists are kind of DEPENDENT
  }

scores_cor_iter <- tibble(seed = 1:100) %>% 
 mutate(scores = map(seed, ~create_scores_cor(seed = .x)))
```

### Simulate Chemical Exposures

Multiple scores and vectors, and add to get the mixture.

```{r}
sim_cor0 <- full_join(scores_cor_iter, patterns_cor_iter, by = "seed")

sim_cor0 <- sim_cor0 %>% 
  mutate(chem = map2(scores, patterns, `%*%`))
```

### Simulate Noise

```{r}
sim_cor <- sim_cor0 %>% 
 mutate(sim = map2(seed, chem, ~add_noise(seed = .x, chem = .y)[[1]]),
        noise = map2(seed, chem, function(x, y) {add_noise(x, y)[[2]]}),
        sim = map(sim, function(x) scale(x, center = FALSE, scale = apply(x, 2, sd))))
```  

```{r}
cormat2 <- round(cor(sim_cor[2,5][[1]][[1]]),2)

melted_cormat2 <- melt(cormat2) %>% rename(Correlation = value)

labelled2 <- melted_cormat2 %>% mutate(Class1 = ifelse(str_detect(Var1, "BDE"), "PBDEs",
                                                     ifelse(str_detect(Var1, "^M"), "Phthalates",
                                                            ifelse(str_detect(Var1, "pcb"), "PCBs",
                                                                   "Phenols")))) %>%
  mutate(Class2 = ifelse(str_detect(Var2, "BDE"), "PBDEs",
                         ifelse(str_detect(Var2, "^M"), "Phthalates",
                                ifelse(str_detect(Var2, "pcb"), "PCBs",
                                       "Phenols")))) %>%
  mutate(Class2 = factor(Class2, levels = c("Phthalates", "Phenols",
                                            "PCBs", "PBDEs")))

ggplot(data = labelled2, aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D",
                       midpoint = 0,
                       na.value = "transparent", limits = c(-1, 1)) +
  theme_grey() + labs(x = "", y = "") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.text = element_blank(),
        panel.spacing = unit(0, "lines"),
        strip.background = element_blank(),
        strip.placement = "outside", legend.position = "bottom") + ggtitle("Overlapping Patterns & Correlated Scores")
```

### Save

```{r}
sim_cor  <- sim_cor %>% rename(true_patterns = patterns, true_scores = scores) %>% dplyr::select(-chem, -noise)
#save(sim_cor, file = "./R/Sims/Iterate/sim_cor.RDA")

# for (i in 1:100) {
#   write_csv(as_tibble(sim_cor[i,4][[1]][[1]]), paste0("./R/Sims/Iterate/sim_cor_", i, "_stand.csv"))
# }
```

## 3: Overlapping Patterns

### Simulate Patterns

Use correlated patterns.

### Simulate Scores

Use distinct scores.

### Simulate Chemical Exposures

Multiple scores and vectors, and add to get the mixture.

```{r}
sim_over0 <- full_join(scores_dist_iter, patterns_cor_iter, by = "seed")

sim_over0 <- sim_over0 %>% 
  mutate(chem = map2(scores, patterns, `%*%`))
```

### Simulate Noise

```{r}
sim_over <- sim_over0 %>% 
 mutate(sim = map2(seed, chem, ~add_noise(seed = .x, chem = .y)[[1]]),
        noise = map2(seed, chem, function(x, y) {add_noise(x, y)[[2]]}),
        sim = map(sim, function(x) scale(x, center = FALSE, scale = apply(x, 2, sd))))
```  

```{r}
cormat3 <- round(cor(sim_over[100,4][[1]][[1]]),2)

summary(sim_over[100,4][[1]][[1]])

melted_cormat3 <- melt(cormat3) %>% rename(Correlation = value)

labelled3 <- melted_cormat3 %>% mutate(Class1 = ifelse(str_detect(Var1, "BDE"), "PBDEs",
                                                     ifelse(str_detect(Var1, "^M"), "Phthalates",
                                                            ifelse(str_detect(Var1, "pcb"), "PCBs",
                                                                   "Phenols")))) %>%
  mutate(Class2 = ifelse(str_detect(Var2, "BDE"), "PBDEs",
                         ifelse(str_detect(Var2, "^M"), "Phthalates",
                                ifelse(str_detect(Var2, "pcb"), "PCBs",
                                       "Phenols")))) %>%
  mutate(Class2 = factor(Class2, levels = c("Phthalates", "Phenols",
                                            "PCBs", "PBDEs")))

ggplot(data = labelled3, aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = Correlation), colour = "white") +
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D",
                       midpoint = 0,
                       na.value = "transparent") +
  theme_grey() + labs(x = "", y = "") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.text = element_blank(),
        panel.spacing = unit(0, "lines"),
        strip.background = element_blank(),
        legend.position = "bottom") + 
        labs(title = "Simulated correlation structure for overlapping patterns & independent scores")
```

### Save

```{r} 
sim_over <- sim_over %>% rename(true_patterns = patterns, true_scores = scores) %>% dplyr::select(-chem, -noise)
#save(sim_over, file = "./R/Sims/Iterate/sim_over.RDA")

# for (i in 1:100) {
#   write_csv(as_tibble(sim_over[i,4][[1]][[1]]), paste0("./R/Sims/Iterate/sim_over_", i, "_stand.csv"))
# }
```

## Test Set

### 1: Distinct Patterns

#### Simulate Patterns

```{r}
patterns_dist_iter_test <- tibble(seed = 101:200) %>% 
 mutate(patterns = map(seed, ~create_patterns_dist(seed = .x)))
```

#### Simulate Scores

```{r}
scores_dist_iter_test <- tibble(seed = 101:200) %>% 
 mutate(scores = map(seed, ~create_scores_dist(seed = .x)))
```

#### Simulate Chemical Exposures

```{r}
sim_dist0_test <- full_join(scores_dist_iter_test, patterns_dist_iter_test, by = "seed")

sim_dist0_test <- sim_dist0_test %>% 
  mutate(chem = map2(scores, patterns, `%*%`))
```

### Simulate Noise

```{r}
sim_dist_test <- sim_dist0_test %>% 
 mutate(sim = map2(seed, chem, ~add_noise(seed = .x, chem = .y)[[1]]),
        noise = map2(seed, chem, function(x, y) {add_noise(x, y)[[2]]}))
```  

### Save

```{r}
sim_dist_test <- sim_dist_test %>% dplyr::select(-noise, -chem)
#save(sim_dist_test, file = "./Sims/Iterate/sim_dist_test.RDA")
```

### 2: Correlated Patterns

#### Simulate Patterns

```{r}
patterns_cor_iter_test <- tibble(seed = 101:200) %>% 
 mutate(patterns = map(seed, ~create_patterns_cor(seed = .x)))
```

#### Simulate Scores

```{r}
scores_cor_iter_test <- tibble(seed = 101:200) %>% 
 mutate(scores = map(seed, ~create_scores_cor(seed = .x)))
```

#### Simulate Chemical Exposures

```{r}
sim_cor0_test <- full_join(scores_cor_iter_test, patterns_cor_iter_test, by = "seed")

sim_cor0_test <- sim_cor0_test %>% 
  mutate(chem = map2(scores, patterns, `%*%`))
```

#### Simulate Noise

```{r}
sim_cor_test <- sim_cor0_test %>% 
 mutate(sim = map2(seed, chem, ~add_noise(seed = .x, chem = .y)[[1]]),
        noise = map2(seed, chem, function(x, y) {add_noise(x, y)[[2]]})) 
```  

#### Save

```{r}
sim_cor_test <- sim_cor_test %>% dplyr::select(-noise, -chem)
#save(sim_cor_test, file = "./Sims/Iterate/sim_cor_test.RDA")
```

### 3: Overlapping Patterns

#### Simulate Patterns

Use correlated patterns.

#### Simulate Scores

Use distinct scores.

#### Simulate Chemical Exposures

```{r}
sim_over0_test <- full_join(scores_dist_iter_test, patterns_cor_iter_test, by = "seed")

sim_over0_test <- sim_over0_test %>% 
  mutate(chem = map2(scores, patterns, `%*%`))
```

#### Simulate Noise

```{r}
sim_over_test <- sim_over0_test %>% 
 mutate(sim = map2(seed, chem, ~add_noise(seed = .x, chem = .y)[[1]]),
        noise = map2(seed, chem, function(x, y) {add_noise(x, y)[[2]]}))
```  

### Save

```{r}
sim_over_test <- sim_over_test %>% dplyr::select(-noise, -chem)
#save(sim_over_test, file = "./Sims/Iterate/sim_over_test.RDA")
```
