---
title: "Loop Simulations"
author: "Lizzy Gibson"
date: "3/5/2020"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
options(scipen = 999)
library(reshape2) # melt
library(gridExtra) # gridarrange
library(bindata) # multivariate binomial dist
library(NMF) # rmatrix
library(MASS) # mvrnorm
library(tidyverse)
library(R.matlab)
```

## 1: Distinct Patterns

### Simulate Patterns

Pattern matrix is 4*50, patterns by chemicals.

```{r}
## Construct a matrix for 4 TRUE exposure patterns
create_patterns_dist <- function (seed) {
  set.seed(seed)
  patterns_dist <- rbind(c(runif(22, 1, 2), rep(0, 11), rep(0, 17)),
                       c(rep(0, 22), runif(11, 1, 2), rep(0, 17)),
                       c(rep(0, 33), runif(9, 1, 2), rep(0, 8)),
                       c(rep(0, 33), rep(0, 9), runif(8, 1, 2)))
  patterns_dist
}

patterns_dist_iter <- tibble(seed = 1:100) %>% 
 mutate(true_patterns = map(seed, create_patterns_dist))
```

### Simulate Scores

```{r}
create_scores_dist <- function (seed) {
  scores_dist <- matrix(nrow = 1000, ncol = 4)
  set.seed(seed)
  scores_dist <- exp(mvrnorm(n = 1000, mu = rep(1, 4), 
                          Sigma = diag(rep(1, 4)))) # Independent log normal dist
  scores_dist
}

scores_dist_iter <- patterns_dist_iter %>% 
 mutate(true_scores = map(seed, create_scores_dist))
```

### Simulate Chemical Exposures

Multiple scores and vectors, and add to get the mixture.

```{r}
sim_dist <- scores_dist_iter %>% 
  mutate(chem = map2(true_scores, true_patterns, `%*%`))
```

### Simulate Noise

```{r}
add_noise <- function (seed, chem) {
  noise <- list(mean=0, sd=1)
  set.seed(seed)
  chem_dist <- pmax(chem + rmatrix(1000, 50, dist=rnorm, mean=noise$mean, sd=noise$sd), 0)	
  chem_dist
}

sim_dist <- sim_dist %>% 
 mutate(sim = map2(seed, chem, add_noise))
```  

## 2: Correlated Patterns

### Simulate Patterns
```{r}
## Construct a matrix for 4 TRUE exposure patterns
create_patterns_cor <- function (seed) {
  set.seed(seed)
  patterns_cor <- rbind(c(runif(22, 1, 2),   runif(11, 0.5, 1), rep(0, 17)),
                         c(rep(0, 22),        runif(11, 1, 2),   runif(9, 0.5, 1),  rep(0, 8)),
                         c(rep(0, 33),                           runif(9, 1, 2),    runif(8, 0.5, 1)),
                         c(runif(22, 0.5, 1), rep(0, 11),        rep(0, 9),         runif(8, 1, 2)))
  patterns_cor
}

patterns_cor_iter <- tibble(seed = 1:100) %>% 
 mutate(true_patterns = map(seed, create_patterns_cor))
```

### Simulate Scores

```{r}
create_scores_cor <- function (seed) {
  scores_cor <- matrix(nrow = 1000, ncol = 4)
  cov <- matrix( c(1,    0.25, 0.15, 0.15,
                  0.25, 1,    0.15, 0.15,
                  0.15, 0.15, 1,    0.25,
                  0.15, 0.15, 0.25, 1), nrow = 4)
  
  set.seed(seed)
  scores_cor <- exp(mvrnorm(n = 1000, mu = rep(1, 4), 
                          Sigma = cov)) # dists are kind of DEPENDENT
  }

scores_cor_iter <- patterns_cor_iter %>% 
 mutate(true_scores = map(seed, create_scores_cor))
```

### Simulate Chemical Exposures

Multiple scores and vectors, and add to get the mixture.

```{r}
sim_cor <- scores_cor_iter %>% 
  mutate(chem = map2(true_scores, true_patterns, `%*%`))
```

### Simulate Noise

```{r}
sim_cor <- sim_cor %>% 
 mutate(sim = map2(seed, chem, add_noise))
```  

## 3: Overlapping Patterns

### Simulate Patterns

Use correlated patterns.

### Simulate Scores

Use distinct scores.

### Simulate Chemical Exposures

Multiple scores and vectors, and add to get the mixture.

```{r}
sim_over <- scores_dist_iter %>% 
  select(seed, true_scores) %>% 
  full_join(., scores_cor_iter %>% select(seed, true_patterns), by = "seed") %>% 
  mutate(chem = map2(true_scores, true_patterns, `%*%`))
```

### Simulate Noise

```{r}
sim_over <- sim_over %>% 
 mutate(sim = map2(seed, chem, add_noise))
```  

## Save

```{r} 
sim_old <- rbind(sim_dist %>% mutate(data = "Distinct"),
      sim_over %>% mutate(data = "Overlapping"),
      sim_cor %>% mutate(data = "Correlated")) %>% 
  mutate(stand = map(sim, function(x) apply(x, 2, function(x) x/sd(x))), #standardize
         var10 = map(stand, function(x) x*sqrt(10))) %>% # scale up var = 10
  pivot_longer(cols = c(sim, stand, var10),
               values_to = "sim",
               names_to = "var") %>% 
  mutate(sim_type = "Raw") %>% 
  mutate(var = ifelse(var == "sim", "reg", var))
  
sim_old

# sim_bnmf <- full_join(sim_old, sim_norm)
#save(sim_old, file = "./R/Sims/Iterate/sim_old.RDA")
save(sim_bnmf, file = "./R/Sims/Iterate/sim_bnmf.RDA")

# for (i in 1:100) {
#   write_csv(as_tibble(sim_over[i,4][[1]][[1]]), paste0("./R/Sims/Iterate/sim_over_", i, "_stand.csv"))
# }
```
